Analyzing Performance Bottleneck in Graph Neural Network Training: An Experimental View

### 2020-08-22 Manuscript/ms.tex
更新修改版的论文.

	\title{Analyzing Performance Bottleneck in Graph Neural Network Training: An Experimental View}
	\author{Zhaokang Wang, Yunpan Wang}%, Chunfeng Yuan, Yihua Huang\corref{correspondingauthor}}
	%	\cortext[correspondingauthor]{Corresponding author}
	\ead{\{wangzhaokang, wangyp\}@smail.nju.edu.cn}%}, \{cfyuan, yhuang\}@nju.edu.cn}
	\address{State Key Laboratory for Novel Software Technology, \\Department of Computer Science and Technology, Nanjing University, \\Nanjing 210023, China}

>>>>

	\title{Analyzing Performance Bottleneck in Graph Neural Network Training: An Experimental View}
    \author{Zhaokang Wang, Yunpan Wang, Chunfeng Yuan, Rong Gu$^*$, Yihua Huang \corref{correspondingauthor}}
    \cortext[correspondingauthor]{Corresponding authors with equal contribution}
    \ead{\{wangzhaokang, wangyp\}@smail.nju.edu.cn, \{cfyuan, gurong, yhuang\}@nju.edu.cn}
	\address{State Key Laboratory for Novel Software Technology, \\Department of Computer Science and Technology, Nanjing University, \\Nanjing 210023, China}

### 2020-08-22 Manuscript/sec7-conclusion.tex
update section 7.
Update section 1.
更新修改版的论文.

\section{Conclusion}

In this work, we systematically study the performance bottleneck in graph neural network training.
We model the existing GNNs with the message-passing framework.
We classify the GNNs according to their edge and vertex calculation complexities to pick four typical GNNs to evaluate.
The experimental results validate our complexity analysis.
The training time and the memory usage increase linearly with the hyper-parameters of a GNN.
To find out the performance bottleneck in time, we decompose the training time per epoch on different levels.
The training time breakdown indicates that the edge calculation and its related basic operators are the performance bottleneck for most GNNs.
Moreover, the intermediate results produced during the edge calculation cause high memory usage, limiting the data scalability.
Adopting sampling techniques can reduce training time and memory usage significantly.
However, the current implementation brings considerable sampling overheads.
The small sampled graphs cannot make full use of the computing power of a GPU either.
Our analysis indicates that the edge calculation should be the main target of optimizations.
Reducing its memory usage and improving its efficiency can significantly improve the performance of GNN training.
Based on the analysis, we propose several potential optimizations for the GNN learning frameworks.
We believe that our analysis can help developers to improving the implementation of the frameworks.

>>>>

\section{Conclusion}
\label{sec:conclusion}

In this work, we systematically study the performance bottleneck in graph neural network training.
We model the existing GNNs with the message-passing framework.
We classify the GNNs according to their edge and vertex calculation complexities to pick four typical GNNs to evaluate.
The experimental results validate our complexity analysis.
The training time and the memory usage increase linearly with the hyper-parameters of a GNN.
To find out the performance bottleneck in time, we decompose the training time per epoch on different levels.
The training time breakdown indicates that the edge calculation and its related basic operators are the performance bottleneck for most GNNs.
Moreover, the intermediate results produced during the edge calculation cause high memory usage, limiting the data scalability.
Adopting sampling techniques can reduce training time and memory usage significantly.
However, the current implementation brings considerable sampling overheads.
The small sampled graphs cannot make full use of the computing power of a GPU either.
Our analysis indicates that the edge calculation should be the main target of optimizations.
Reducing its memory usage and improving its efficiency can significantly improve the performance of GNN training.
Based on the analysis, we propose several potential optimizations for the GNN learning frameworks.
We believe that our analysis can help developers to have a better understanding of the characteristics of GNN computation.

\section*{Acknowledgement}

This work is funded in part by National Key R\&D Program of China [grant number 2019YFC1711000]; China NSF Grants [grant number U1811461]; Jiangsu Province Industry Support Program [grant number BE2017155]; Natural Science Foundation of Jiangsu Province [grant number BK20170651]; Collaborative Innovation Center of Novel Software Technology and Industrialization; and the program B for Outstanding PhD candidate of Nanjing University.
